---
title: "ExtDataFig4D"
author: "N. Alcala"
date: "`r Sys.Date()`"
output: html_document
---

# Figure 4: Supra-carcinoids

This document produces Figure 4 from the lungNENomics manuscript, describing the characteristics of supra-carcinoids.

## Panel A

### Load libraries

```{r pressure, echo=FALSE}
library(tidyverse)
library(patchwork)
library(ggbeeswarm)
```

### load lungNENomics color schemes
For now, this only runs locally. Nevertheless, not necessary to run, just provides cosmetics.
```{r}
source("/data/lungNENomics/work/Descriptive_manuscript_data/Colour_palettes.R")
colors_clusters = c("Ca A1"="#999933", "Ca A2"= "#DDCC77", "Ca B"= "#117733", "sc-enriched"="#CC6677",  
                    "Typical"="#9DB802","Atypical"="#025B0E","undetermined"="#CEDB80",
                    "LCNEC"="#824833" ,"SCLC"="#000000",
                    SINET = "#2ad4ffff") #, "Adeno-Ca"="#ff80e5ff", "Squamous-Cell-Ca"="#8080ffff") 
```


### load metadata
We need to merge the metadata for the LCNEC and SCLC cohorts (data harmonized in Gabriel et al. 2020) with the clinical data from the lungNENomics cohort (in Table S1)

```{r}
Attributes = read_tsv(unzip('../../Data/GabrielEtAl2020/Attributes.txt.zip')[1])

load("/data/lungNENomics/work/Descriptive_manuscript_data/MOFA/MOFA.Exp.Meth.Alt.CNV_lungNENomicsCombined/ParetoTI/variables_archetypes_MOFA_LNET.RData") # data also contained in Table S1
gigascience_attributes = read_tsv("/data/lungNENomics/work/SextonoatesA/Attributes.txt")
var_LNET$sample_id = as.character(var_LNET$sample_id)
var_LNET$group = var_LNET$archetype_k4_LF3_label

var_cli = bind_rows(var_LNET, Attributes %>% filter(Histopathology_simplified%in%c("LCNEC","SCLC")) %>% 
                      mutate(sample_id = Sample_ID, group=Histopathology_simplified) %>% dplyr::select(sample_id,group) )
var_cli = bind_rows(var_cli,tibble(sample_id=c("LCNEC3T","LCNEC4T"),group="LCNEC") )
var_cli$type[is.na(var_cli$type)] = var_cli$group[is.na(var_cli$type)]

supraca_table = readxl::read_xlsx("../../Data/Supplementary_Table_S36.xlsx",skip=1,n_max = 19)
```


### Load classification data
```{r}
supraca_table = supraca_table %>% dplyr::select(manuscript_id,consensus_pathology,asbestos,driver_mutations,archetype,`supra-carcinoid`)
for(i in 1:ncol(supraca_table)) supraca_table[[i]] = as.factor(supraca_table[[i]])

supraca_table.long = supraca_table %>% pivot_longer(cols = c(consensus_pathology,asbestos,driver_mutations,archetype,`supra-carcinoid`))
```

We reorder the variables
```{r}
supraca_table.long = supraca_table.long %>% mutate(value=factor(value,levels=c( "Typical","Carcinoid","NA","Atypical","NET G3",
                                                           "no","yes",
                                                           "ARID1A (nonsense); BRAF (missense)", "BRAF (missense)", 
                                                           "ATM (nonsense); MEN1 (nonsense)", "MEN1 (frame shift indel)", 
                                                           "Ca A1", "sc-enriched", 
                                                           "FALSE", "TRUE" )), 
                              name=factor(name,levels=rev(c("archetype","supra-carcinoid","asbestos","consensus_pathology", "driver_mutations"))) )
```

### Plots
```{r}
cols_4A = c(type5[1],type5[3],"NA"="white",type5[2],type5[4],
            "no","yes",
            "ARID1A (nonsense); BRAF (missense)", "BRAF (missense)","ATM (nonsense); MEN1 (nonsense)", "MEN1 (frame shift indel)", 
            arc4[1], arc4[4],
            "FALSE", "TRUE")

Fig4A <- ggplot(data=supraca_table.long , 
             aes(x=manuscript_id, y = name,fill=value )) + geom_tile(col="white") +
  theme_classic() + theme(axis.text.x=element_text(angle = 90,hjust = 1,vjust=0.5))
  #geom_violin(aes(fill=archetype_k4_LF3_label)) + 
  #geom_boxplot(fill="white", width=0.1, outlier.size = 0.4) +
  #scale_fill_manual(name = "Archetype", values=c("Ca A1"="#999933", "Ca A2"="#DDCC77", "Ca B"="#117733", "sc-enriched"="#CC6677")) +
  #theme_classic() + ylab("MKI67 expression [log10(TPM+1)]") + xlab("Archetype") +
  #theme(legend.position = "right") +
  #scale_y_continuous(breaks = c(0,0.5,1,1.5,2), limits=c(0,2.5)) +
  #guides(color=guide_legend(nrow=1,byrow=TRUE)) + 
  #geom_segment(aes(x=1,xend=2,y=ylabpos,yend=ylabpos)) +  geom_text(aes(label = "*", y = ylabpos+0.1, x=1.5)) +
  #geom_segment(aes(x=1,xend=4,y=ylabpos+.3,yend=ylabpos+0.3)) +  geom_text(aes(label = "***", y = ylabpos+0.3+0.1, x=2.5)) +
  #geom_segment(aes(x=2,xend=4,y=ylabpos+0.3*2,yend=ylabpos+0.3*2)) +  geom_text(aes(label = "***", y = ylabpos+0.3*2+0.1, x=3)) +
  #geom_segment(aes(x=3,xend=4,y=ylabpos+0.3*3,yend=ylabpos+0.3*3)) +  geom_text(aes(label = "***", y = ylabpos+0.3*3+0.1, x=3.5)) +
  #theme(axis.text.y=element_text(size=11), axis.text.x=element_text(size=11), 
  #      axis.title.x=element_text(size=11), axis.title.y=element_text(size=11)) + 
  #theme(legend.text=element_text(size=11), legend.title=element_text(size=11)) +
  #labs(title="MKI67 expression by group") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Molecular group") + guides(fill="none")

ggsave("/data/lungNENomics/work/alcalan/Figures/ExtendedDataFig_4D_MKI67.svg",plot = EDF4Dleft,height=2.5,width=2.5)
```