---
title: "ExtendedDataFig3A"
author: "Laurane Mangé"
date: "2025-03-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document produces pannel A from Extended Data Figure 3 from the lungNENomics manuscript, displaying the number of hallmarks of cancer (<https://aacrjournals.org/cancerdiscovery/article/12/1/31/675608/Hallmarks-of-Cancer-New-DimensionsHallmarks-of>) affected by damaging mutations per patient for each archetype. All file paths used are to be modified according to the user's working directory.

## Libraries

The first step of the study is to load the required libraries.

```{r libraries}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(tidyr)
library(forestmodel)
library(rlang)
library(eulerr)
library(cowplot)
library(patchwork)
library(writexl)
library(jtools)
```

## Fetching the necessary data

We now load the necessary data for this study : Samples displaying genes with damaging SNV mutations, as well as archetypes for each sample and hallmarks of cancer affected through mutation for each gene.

```{r fetch data}
# dataframe with hallmarks from COSMIC
clean_hallmarks <- read_csv("/home/mangel/Hallmarks/Stage IARC/clean_hallmarks_COSMICv99_Intogen2023-05-31.tsv")

clean_hallmarks <- clean_hallmarks %>% dplyr::select(-c(1,11,12,13))

all_hm <- sort(unique(clean_hallmarks$HALLMARK))

# Pareto archetype color palette
arc4 <- c("Ca A1"="#999933", "Ca A2"="#DDCC77", "Ca B"="#117733", "sc-enriched"="#CC6677", "LCNEC"="#882255", "SCLC"="#332288") 
### Hallmark profile colors
hmPalette <- c("#CC0000", "#006666", "#330066", "#CC2266", "#666666", "#003399", "#111111", "#336600", "#663333", "#CC6600")

# dataframe with damaging SNVs
snv_dmg <- get(load('dataset_snv_dmg_lungNENomicsCombined.RData'))
unique(snv_dmg$gene1_name)
unique(snv_dmg$sample_id) #102 samples

dmg_genes_list <- unique(snv_dmg$gene1_name[which(snv_dmg$gene1_name %in% clean_hallmarks$GENE_SYMBOL)])
# 53 genes from clean_hallmarks (COSMIC)

# Making snv_dmg_arch : dataframe with damaging SNVs and archetypes per sample
variables_archetypes_MOFA_LNET <- get(load('variables_archetypes_MOFA_LNET.RData')) #319 samples

snv_dmg_arch <- get(load('variables_archetypes_MOFA_LNET.RData'))
snv_dmg_arch <- merge(snv_dmg_arch, snv_dmg[c('sample_id','gene1_name')], by = 'sample_id')
snv_dmg_arch <- snv_dmg_arch %>% relocate(gene1_name, .before=Factor1)
snv_dmg_arch <- snv_dmg_arch %>% relocate(archetype_k4_LF3_label, .before=Factor1)
snv_dmg_arch <- merge(snv_dmg_arch, clean_hallmarks[c('GENE_SYMBOL','HALLMARK')], by.x = 'gene1_name', by.y = 'GENE_SYMBOL',all=TRUE)
snv_dmg_arch <- snv_dmg_arch %>% relocate(HALLMARK, .before = Factor1)
snv_dmg_arch <- snv_dmg_arch %>% dplyr::filter(!is.na(sample_id))
snv_dmg_arch <- snv_dmg_arch %>% dplyr::filter(!duplicated(snv_dmg_arch))
snv_dmg_arch <- snv_dmg_arch %>% rename(archetype = archetype_k4_LF3_label)
unique(snv_dmg_arch$sample_id) # 102 good

```

```{r 53 genes}
table(snv_dmg_arch$archetype)
unique(snv_dmg_arch %>% filter(!is.na(HALLMARK)) %>% pull(gene1_name)) # 53 dmg genes with hm affected
table(snv_dmg_arch$HALLMARK)
```
There are 53 genes with damaging SNV mutations that affect at least one of the 10 known hallmarks of cancer.

We now create dataframes in "wide" format giving information about whether a hallmark is affected (1) or not (0) by the mutation of a given gene (first column).

```{r gene_hm.wide}
gene_hm.wide = snv_dmg_arch[c('gene1_name','HALLMARK')] %>% filter(!duplicated(snv_dmg_arch[c('gene1_name','HALLMARK')])) %>% mutate(Presence=1) %>% 
  pivot_wider(id_cols = gene1_name, names_from = HALLMARK,values_from = Presence,values_fill = 0,names_expand = TRUE)

Missing <- setdiff(all_hm, names(gene_hm.wide[-1]))
gene_hm.wide[Missing] <- 0
gene_hm.wide <- gene_hm.wide[c('gene1_name',sort(colnames(gene_hm.wide[-1])))]
gene_hm.wide <- gene_hm.wide %>% select(-c("NA"))

```

We can make the same dataframe for archetypes instead of genes.

```{r arch_hm.wide}
arch_hm.wide = snv_dmg_arch[c('archetype','HALLMARK')] %>% filter(!duplicated(snv_dmg_arch[c('archetype','HALLMARK')])) %>% mutate(Presence=1) %>% 
  pivot_wider(id_cols = archetype, names_from = HALLMARK,values_from = Presence,values_fill = 0,names_expand = TRUE)

Missing <- setdiff(all_hm, names(arch_hm.wide[-1]))
arch_hm.wide[Missing] <- 0
arch_hm.wide <- arch_hm.wide[c('archetype',sort(colnames(arch_hm.wide[-1])))]
arch_hm.wide <- arch_hm.wide %>% select(-c("NA"))
```

All 4 archetypes seem to have (almost) all hallmarks affected.

However, the dataframe in "wide" format we will use the most contains the previously mentioned hallmark information per sample (i.e. per patient).

```{r sample_hm.wide}
table(snv_dmg_arch$archetype,snv_dmg_arch$sample_id)
length(unique(snv_dmg_arch$sample_id)) # 102 different samples

sample_hm.wide = snv_dmg_arch[c('sample_id','HALLMARK')] %>% filter(!duplicated(snv_dmg_arch[c('sample_id','HALLMARK')])) %>% mutate(Presence=1) %>% 
  pivot_wider(id_cols = sample_id, names_from = HALLMARK,values_from = Presence,values_fill = 0,names_expand = TRUE)

Missing <- setdiff(all_hm, names(sample_hm.wide[-1]))
sample_hm.wide[Missing] <- 0
sample_hm.wide <- sample_hm.wide[c('sample_id',sort(colnames(sample_hm.wide[-1])))]

# samples belonging to each archetype
sc_samples <- unique(snv_dmg_arch %>% filter(archetype=="sc-enriched") %>% pull(sample_id))
a1_samples <- unique(snv_dmg_arch %>% filter(archetype=="Ca A1") %>% pull(sample_id))
a2_samples <- unique(snv_dmg_arch %>% filter(archetype=="Ca A2") %>% pull(sample_id))
b_samples <- unique(snv_dmg_arch %>% filter(archetype=="Ca B") %>% pull(sample_id))

### fisher tests on hallmarks within archetypes

fish_arch_hm  <- data.frame(matrix(nrow = 0, ncol = 4)) 
names(fish_arch_hm) <- c('archetype','hm','fisherTestPvalue','fisherTestOddsRatio')

for (i in 1:length(unique(snv_dmg_arch$archetype))) {
  for(j in 1:length(unique(snv_dmg_arch$HALLMARK)[-1])){
    fish_arch_hm[nrow(fish_arch_hm)+1, ] <- c(unique(snv_dmg_arch$archetype)[i],unique(snv_dmg_arch$HALLMARK)[j+1],round(fisher.test(table(snv_dmg_arch$archetype==unique(snv_dmg_arch$archetype)[i],snv_dmg_arch$HALLMARK==unique(snv_dmg_arch$HALLMARK)[j+1]))$p.value,3),round(fisher.test(table(snv_dmg_arch$archetype==unique(snv_dmg_arch$archetype)[i],snv_dmg_arch$HALLMARK==unique(snv_dmg_arch$HALLMARK)[j+1]))$estimate,3))
  }
}

fish_arch_hm
fish_arch_hm[which(fish_arch_hm$fisherTestPvalue<=0.05),]
# signif : Ca A1 angiogenesis more than expected and Ca A2 suppression of growth same

```

Now we plot the distribution of the number of hallmarks affected per patients for each archetype in the form of a violin plot.

```{r violin}
sample_hm_arch.wide <- merge(sample_hm.wide,snv_dmg_arch[c("sample_id","archetype")],by="sample_id")
sample_hm_arch.wide <- sample_hm_arch.wide %>% filter(!duplicated(sample_hm_arch.wide))

sample_sizes <- as.data.frame(cbind(unique(sample_hm_arch.wide$archetype),c(length(a1_samples),length(a2_samples),length(b_samples),length(sc_samples))))
colnames(sample_sizes) <- c("archetype","n")
sample_sizes$n <- as.numeric(sample_sizes$n)

wilcox.test(rowSums(sample_hm_arch.wide %>% filter(archetype=="Ca A1") %>% select(-c(1,9,13))),rowSums(sample_hm_arch.wide %>% filter(archetype=="Ca A2") %>% select(-c(1,9,13))),paired=FALSE)
wilcox.test(rowSums(sample_hm_arch.wide %>% filter(archetype=="Ca A1") %>% select(-c(1,9,13))),rowSums(sample_hm_arch.wide %>% filter(archetype=="Ca B") %>% select(-c(1,9,13))),paired=FALSE)
wilcox.test(rowSums(sample_hm_arch.wide %>% filter(archetype=="Ca A1") %>% select(-c(1,9,13))),rowSums(sample_hm_arch.wide %>% filter(archetype=="sc-enriched") %>% select(-c(1,9,13))),paired=FALSE)
wilcox.test(rowSums(sample_hm_arch.wide %>% filter(archetype=="Ca A2") %>% select(-c(1,9,13))),rowSums(sample_hm_arch.wide %>% filter(archetype=="Ca B") %>% select(-c(1,9,13))),paired=FALSE)
wilcox.test(rowSums(sample_hm_arch.wide %>% filter(archetype=="Ca A2") %>% select(-c(1,9,13))),rowSums(sample_hm_arch.wide %>% filter(archetype=="sc-enriched") %>% select(-c(1,9,13))),paired=FALSE)
wilcox.test(rowSums(sample_hm_arch.wide %>% filter(archetype=="Ca B") %>% select(-c(1,9,13))),rowSums(sample_hm_arch.wide %>% filter(archetype=="sc-enriched") %>% select(-c(1,9,13))),paired=FALSE)
# sc-enriched diff with every other group

#my_comparisons <- list( c("Ca A1", "Ca A2"), c("Ca A1", "Ca B"), c("Ca A1", "sc-enriched"), c("Ca A2", "Ca B"), c("Ca A2", "sc-enriched"), c("Ca B", "sc-enriched"))
my_comparisons <- list(c("Ca A1", "sc-enriched"), c("Ca A2", "sc-enriched"), c("Ca B", "sc-enriched"))
#only these are significant

violin <- ggplot(sample_hm_arch.wide, aes(x=archetype, y=rowSums(sample_hm_arch.wide[,c(2:8,10:12)]), fill=archetype, color=archetype)) + 
  geom_violin() +
  geom_boxplot(width=0.15, color="black", fill="white") +
  geom_dotplot(data=sample_hm_arch.wide %>% filter(archetype=="sc-enriched"),aes(x=archetype,y=rowSums(sample_hm_arch.wide %>% filter(archetype=="sc-enriched") %>% select(c(2:8,10:12)))), binaxis = "y", stackdir='center', color="white") +
  scale_color_manual(values=arc4) +
  scale_fill_manual(values=arc4) +
  scale_y_continuous(breaks = seq(0, 10, by = 1)) +
  labs(y="Number of hallmarks affected per sample") +
  geom_text(data = sample_sizes,aes(x=archetype,y=-1,label=n)) +
  theme_bw() +
  theme(panel.grid=element_blank(), panel.border = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust=1,colour =arc4[(c("Ca A1","Ca A2","Ca B","sc-enriched"))]))
violin <- violin + stat_compare_means(comparisons = my_comparisons, method = "wilcox.test", label = "p.signif")
violin

```