---
title: "Figure 2b"
author: "Emilie Matthian, Nicolas Alcala, Alexandra Sexton-Oates"
date: "12/08/2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

This R markdown produces Figure 2b, Extended Data Figure 2c and Extended Data Figure 2e 

### Load libraries 
```{r}
library(openxlsx)
library(tidyr)
library(dplyr)
library(ggplot2)
library(patchwork)
```

### Load and format datasets required
```{r}
Table_S1 <- read.xlsx("Supplementary_Table_S1.xlsx", startRow = 4) # sample information 
samples <- Table_S1$sample.id[which(Table_S1$cohort.LNET=="TRUE" & grepl("WGS", Table_S1$omics.group))] # list of relevant samples

Table_S15 <- read.xlsx("Supplementary_Table_S15.xlsx", startRow = 3) # damaging SNVs
all(Table_S15$sample_id %in% Table_S1$sample.id) 
Table_S15 <- Table_S15[which(Table_S15$sample_id %in% samples), ] # filter for relevant samples

Table_S16 <- read.xlsx("Supplementary_Table_S16.xlsx", startRow = 3) # damaging structural variants
all(Table_S16$sample_id %in% Table_S1$sample.id) 
Table_S16 <- Table_S16[which(Table_S16$sample_id %in% samples), ] # filter for relevant samples

Table_S23 <- read.xlsx("Supplementary_Table_S23.xlsx", startRow = 3) # driver genes
Table_S23 <- Table_S23[c(1:13), ] # keep only required rows

gistic <- read.xlsx("gistic_table.xlsx") # Gistic CNV event data - currently in One Drive
all(gistic$sample_id %in% Table_S1$sample.id) 
gistic <- gistic[which(gistic$sample_id %in% samples), ] # filter for relevant samples

Table_S17 <- read.xlsx("Supplementary_Table_S17.xlsx", startRow = 3) # WGD and shattered region information 
Table_S17 <- Table_S17[c(1:145), ] # only require top part of the table 
Table_S17 <- Table_S17[which(Table_S17$sample_id %in% samples), ] # filter for relevant samples

Table_S24 <- read.xlsx("Supplementary_Table_S24.xlsx", startRow = 3) # List of epigenetic regulatory genes 
Table_S24 <- Table_S24[which(Table_S24$TPM.filter.pass=="YES"), ] # keep only genes which pass TPM filter 
```


### Build oncoplot dataframe 
SNV and SV in driver genes
```{r}
drivers <- Table_S23$gene_name[which(Table_S23$`Pass.expression.filter.(TPM>1.in.≥80%.of.cohort)`=="YES")] # drivers which pass TPM expression filter

any(drivers %in% Table_S15$gene1_name) # TRUE
any(drivers %in% Table_S15$gene2_name) # FALSE

any(drivers %in% Table_S16$gene1_name) # FALSE
any(drivers %in% Table_S16$gene2_name) # FALSE

dataset_snv_driver <- Table_S15[which(Table_S15$gene1_name %in% drivers),]

# Check if any drivers have multiple hits 
genes_frequency <- as.data.frame(sort(unique(dataset_snv_driver$gene1_name)))
colnames(genes_frequency) <- "gene1_name"
genes_frequency$samples <- sapply(genes_frequency$gene1_name, function(x) paste0(dataset_snv_driver$sample_id[which(dataset_snv_driver$gene1_name==x)], collapse = ";"))
genes_frequency$length_all <- sapply(genes_frequency$samples, function(x) length(unlist(strsplit(x, ";"))))
genes_frequency$length_unique <- sapply(genes_frequency$samples, function(x) length(unique(unlist(strsplit(x, ";")))))
genes_frequency$gene1_name[which(genes_frequency$length_all!=genes_frequency$length_unique)] #  LNEN052_TU has two SNVs in CLIP1

dataset_snv_driver[which(dataset_snv_driver$gene1_name == "CLIP1"),c(1,12,23)]
dataset_snv_driver[c("74","75"),23] <- "Multi Hit" # change variant label of CLIP1 alterations in LNEN052_TU to multi hit

# Make oncoplot dataframe 
oncoplot_temp <- dataset_snv_driver[,c(1,29,23)]
oncoplot_temp <- oncoplot_temp[!(duplicated(oncoplot_temp)),] # Remove rows which correspond to a sample having multiple entries for the same gene (all labelled multi hit)
rownames(oncoplot_temp) <- NULL
oncoplot_temp <- oncoplot_temp %>% pivot_wider(names_from = gene1_name, values_from = Variant_Label)
oncoplot_temp[is.na(oncoplot_temp)] <- "WT"
oncoplot_temp <- as.data.frame(oncoplot_temp)
```

Gistic events
```{r}
dim(gistic) # 102 samples 
dim(oncoplot_temp) # 35 samples 
dim(oncoplot_temp[oncoplot_temp$sample_id %in% gistic$sample_id, ]) # 35 
length(setdiff(gistic$sample_id, oncoplot_temp$sample_id )) # 67 

# Merge datasets 
oncoplot_temp2 <- merge(oncoplot_temp, gistic, by="sample_id", all.x=TRUE, all.y = TRUE)
oncoplot_temp2[is.na(oncoplot_temp2)] <- "WT" # make samples with CNV data but no SNVs WT for SNVs
```

WGD status
```{r}
all(oncoplot_temp2$sample_id %in% Table_S17$sample_id) # TRUE
oncoplot_temp2$WGD <- sapply(oncoplot_temp2$sample_id, function(x) Table_S17$wholeGenomeDuplication[which(Table_S17$sample_id==x)])
oncoplot_temp2$WGD <- ifelse(oncoplot_temp2$WGD=="TRUE", "WGD", "WT")
```

Shattered region status
```{r}
oncoplot_temp2$shattered_regions <- sapply(oncoplot_temp2$sample_id, function(x) Table_S17$shattered_regions_presence[which(Table_S17$sample_id==x)])
oncoplot_temp2$shattered_regions <- ifelse(oncoplot_temp2$shattered_regions=="None", "WT",
                                           ifelse(oncoplot_temp2$shattered_regions=="High-Confidence", "High confidence", "Low confidence")) 
```

SNVs/SVs in epigenetic regulatory genes other than drivers
```{r}
epigenetic_genes <- Table_S24$gene.name
epigenetic_genes <- epigenetic_genes[!(epigenetic_genes %in% drivers)] # exclude ARID1A, ASXL1, ATM and MEN1 as they're already reported in oncoplot
# leaves n = 357

any(Table_S15$gene1_name %in% epigenetic_genes) # TRUE
any(Table_S15$gene2_name %in% epigenetic_genes) # FALSE

any(Table_S16$gene1_name %in% epigenetic_genes) # TRUE 
any(Table_S16$gene2_name %in% epigenetic_genes) # FALSE

dataset_snv_epigenetic <- Table_S15[which(Table_S15$gene1_name %in% epigenetic_genes),c(1,29,23)]
dataset_sv_epigenetic <- Table_S16[which(Table_S16$gene1_name %in% epigenetic_genes),c(1,21,17)]

samples_erg_alt <- c(dataset_snv_epigenetic$sample_id, dataset_sv_epigenetic$sample_id)
samples_erg_alt <- unique(samples_erg_alt) # n = 34 

oncoplot_temp2$ERG <- sapply(oncoplot_temp2$sample_id, function(x) if(x %in% samples_erg_alt){"ERG"}else{"WT"})
```


### Draw oncoplots

#### Extended Data Figure 2c 
```{r, fig.height=12, fig.height=10}
var_cli <- Table_S1[which(Table_S1$cohort.LNET=="TRUE" & grepl("WGS", Table_S1$omics.group)),]
var_cli <- var_cli[,c(1,13,24,26,27,47)]
colnames(var_cli) <- c("sample_id", "type", "sex", "age", "location", "molecular_group")

oncoplot_temp2$group <- sapply(oncoplot_temp2$sample_id, function(x) var_cli$molecular_group[which(var_cli$sample_id==x)])
oncoplot <- oncoplot_temp2
rownames(oncoplot) <- NULL 

# Get the sample order  
oncoplot$group <- factor(oncoplot$group , levels = c("Ca A1", "Ca A2", "Ca B", "sc-enriched"))
oncoplot <- oncoplot[order(oncoplot$group),]
sample_order <- left_join(oncoplot,var_cli) %>% arrange(group, MEN1, ARID1A, EIF1AX, ATM, `HLA-C`, `HLA-A`, ASXL1, BRAF, CLIP1, SGK1, `del_1p13.3`, del_3p, del_3q, amp_5q, amp_5p, amp_7p, `del_11q13.1`, amp_14p, WGD, shattered_regions, ERG, type) %>% pull(sample_id)

# arrange the oncoplot dataframe and set sample order 
oncoplot <- oncoplot[,-ncol(oncoplot)]  
oncoplot <- oncoplot %>% gather(Feature, Value, c(2:ncol(oncoplot))) 
oncoplot$sample_id <- factor(oncoplot$sample_id, levels=sample_order) 

gene_order <- c("MEN1", "ARID1A", "EIF1AX", "ATM", "HLA-C", "HLA-A", "ASXL1", "BRAF", "CLIP1", "SGK1")
gistic_order <- c("del_1p13.3", "del_3p", "del_3q", "amp_5q", "amp_5p", "amp_7p", "del_11q13.1", "amp_14p")
large_genome_event_order <- c("WGD", "shattered_regions", "ERG")
col_order <- c(gene_order, gistic_order, large_genome_event_order)
col_order <- rev(col_order)
oncoplot$Feature <- factor(oncoplot$Feature, levels=col_order) 

# set oncoplot colours and make plot
colors_snv <- c("Frame Shift Indel"="#1965B0", "Missense"="#4EB265", "Multi Hit"="#882E72", "Nonsense"="#DC050C", "Splice Site"="#F7F056", 
                "CNV_event"="#FF9F00", "WT"="white", "WGD"="#C4DE54", "High confidence"="#81031D", "Low confidence"="#4E156F", "ERG"="#4EC6D0") 

p1 <- ggplot(oncoplot, aes(sample_id, Feature, fill=Value)) + 
  geom_tile(color="white") + # previous colour: #E7E7E7
  scale_fill_manual(name = "Features", values = colors_snv, limits = names(colors_snv)) +
  xlab(label = "Samples") + ylab(label = "") + theme_classic() + theme(axis.text.y = element_text(face = "italic")) +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) + theme(legend.position = "bottom") + geom_hline(yintercept = c())

# arrange clinical plot dataframe and set sample order 
clinplot <- var_cli[,c("sample_id","molecular_group","type","sex","age","location")]
clinplot <- clinplot[which(clinplot$sample_id %in% sample_order),]
colnames(clinplot)[2:6] <- c("Molecular group", "Type", "Sex", "Age", "Location")  
clinplot$Sex <- ifelse(clinplot$Sex=="M", "Male","Female")
clinplot$Location <- ifelse(clinplot$Location=="proximal","Proximal", ifelse(clinplot$Location=="distal","Distal",NA))
clinplot <- clinplot %>% gather(Feature, Value, c(2:6)) # Make wide dataframe long for plotting
clinplot$sample_id <- factor(clinplot$sample_id, levels=sample_order) # set the sample order correctly for plotting
clinplot$Feature <- factor(clinplot$Feature, levels=c("Molecular group","Type","Sex","Age","Location"))

# set clinical plot colours and make plot 
colors_clin <- c("Proximal"="#999933", "Distal"="#AA4499", 
                 "16-40"="#feda8b", "41-65"="#f67e4b", "66-90"="#A50026", 
                 "Male"="#364B9A", "Female"="#A50026", 
                 "Typical"="#9DB802","Atypical"="#025B0E","Carcinoid"="#CEDB80",
                 "Ca A1"="#999933", "Ca A2"="#DDCC77", "Ca B"="#117733", "sc-enriched"="#CC6677") 

p2 <- ggplot(clinplot, aes(sample_id, Feature, fill=Value)) + 
  geom_tile(color="white") + # previous colour: #E7E7E7
  scale_fill_manual(name = "Features", values = colors_clin, limits = names(colors_clin), na.value="white") +
  xlab(label = "") + ylab(label = "Sample features") + theme_classic() +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) + theme(legend.position = "top")

# combine clinical plot and oncoplot 
patchwork <- p2/p1 + plot_layout(heights=c(0.3,1)) + plot_annotation(tag_levels = list(c('c'))) &  theme(plot.tag = element_text(size = 20)) 
patchwork
```

#### Figure 2d 
```{r, fig.height=12, fig.height=6}
oncoplot <- oncoplot_temp2 %>% dplyr::select(sample_id, MEN1, ARID1A, EIF1AX, ATM, `HLA-C`, `HLA-A`, ASXL1, BRAF, CLIP1, SGK1, `del_11q13.1`,
                WGD, shattered_regions, ERG, group)
rownames(oncoplot) <- NULL

# Get the sample order 
oncoplot$group <- factor(oncoplot$group , levels = c("Ca A1", "Ca A2", "Ca B", "sc-enriched"))
oncoplot <- oncoplot[order(oncoplot$group),]
sample_order <- left_join(oncoplot,var_cli) %>% arrange(group, MEN1,`del_11q13.1`, ARID1A, EIF1AX, ATM, `HLA-C`, `HLA-A`, ASXL1, BRAF, CLIP1, SGK1, WGD, shattered_regions, ERG, type) %>% pull(sample_id)

# arrange the oncoplot dataframe and set sample order 
oncoplot <- oncoplot[,-ncol(oncoplot)]  
oncoplot <- oncoplot %>% gather(Feature, Value, c(2:ncol(oncoplot))) 
oncoplot$sample_id <- factor(oncoplot$sample_id, levels=sample_order) 

gene_order <- c("MEN1", "del_11q13.1", "ARID1A", "EIF1AX", "ATM", "HLA-C", "HLA-A", "ASXL1", "BRAF", "CLIP1", "SGK1")
large_genome_event_order <- c("WGD", "shattered_regions", "ERG")
col_order <- c(gene_order, large_genome_event_order)
col_order <- rev(col_order)
oncoplot$Feature <- factor(oncoplot$Feature, levels=col_order) 

# set oncoplot colours and make plot
colors_snv <- c("Frame Shift Indel"="#1965B0", "Missense"="#4EB265", "Multi Hit"="#882E72", "Nonsense"="#DC050C", "Splice Site"="#F7F056", 
                "CNV_event"="#FF9F00", "WT"="white", "WGD"="#C4DE54", "High confidence"="#81031D", "Low confidence"="#4E156F", "ERG"="#4EC6D0") 

p3 <- ggplot(oncoplot, aes(sample_id, Feature, fill=Value)) + 
  geom_tile(color="white") + # previous colour: #E7E7E7
  scale_fill_manual(name = "Features", values = colors_snv, limits = names(colors_snv)) +
  xlab(label = "Samples") + ylab(label = "") + theme_classic() + theme(axis.text.y = element_text(face = "italic")) +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) + theme(legend.position = "bottom") + 
  geom_vline(xintercept = c(44,44+29,44+29+24)+0.5,col="black")

# arrange clinical plot dataframe and set sample order 
clinplot <- var_cli[,c("sample_id","molecular_group","type")]
clinplot <- clinplot[which(clinplot$sample_id %in% sample_order),]
colnames(clinplot)[2:3] <- c("Molecular group", "Type")  
clinplot <- clinplot %>% gather(Feature, Value, c(2:3)) # Make wide dataframe long for plotting
clinplot$sample_id <- factor(clinplot$sample_id, levels=sample_order) # set the sample order correctly for plotting
clinplot$Feature <- factor(clinplot$Feature, levels=c("Type","Molecular group"))

# set clinical plot colours and make plot 
colors_clin <- c("Typical"="#9DB802","Atypical"="#025B0E","Carcinoid"="#CEDB80",
                 "Ca A1"="#999933", "Ca A2"="#DDCC77", "Ca B"="#117733", "sc-enriched"="#CC6677") 

p4 <- ggplot(clinplot, aes(sample_id, Feature, fill=Value)) + 
  geom_tile(color="white") + # previous colour: #E7E7E7
  scale_fill_manual(name = "Features", values = colors_clin, limits = names(colors_clin), na.value="white") +
  xlab(label = "") + ylab(label = "Sample features") + theme_classic() +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) + theme(legend.position = "top")

# combine clinical plot and oncoplot 
patchwork <- p4 /p3 + plot_layout(heights=c(0.3,1)) + plot_annotation(tag_levels = list(c('b'))) &  theme(plot.tag = element_text(size = 20)) 
patchwork

# make and plot barplot
oncoplot.barplot <- left_join(oncoplot,var_cli[,c("sample_id","molecular_group","type")]) %>% filter(Value!="WT") %>% group_by(Feature, molecular_group) %>% summarize(n=n())
oncoplot.barplot$Feature <- factor(oncoplot.barplot$Feature, col_order)

ggplot(oncoplot.barplot,aes(y=Feature,x=n/102*100,fill=molecular_group)) + geom_bar(stat="identity",width=0.95,show.legend = F) + 
  scale_fill_manual(values=c("Ca A1"="#999933", "Ca A2"="#DDCC77", "Ca B"="#117733", "sc-enriched"="#CC6677")) + theme_classic() + coord_cartesian(expand = F)
```

Extended Data Figure 2e
```{r}
summary_events <- oncoplot_temp2

summary_events$SNV_event_driver_gene <- ifelse(apply(summary_events[, c(2:11)], 1, function(x) any(x != "WT")), 1, 0)
summary_events$CNV_event <- ifelse(apply(summary_events[, c(12:19)], 1, function(x) any(x != "WT")), 1, 0)
summary_events$WGD_event <- ifelse(summary_events$WGD != "WT", 1, 0)
summary_events$shattered_event <- ifelse(summary_events$shattered_regions != "WT", 1, 0)
summary_events$Epigenetic_regulatory_gene_event <- ifelse(summary_events$ERG != "WT", 1, 0)

summary_events <- summary_events[,c(1,23:28)]
summary_events$several_events <- ifelse(apply(summary_events[, c(3:7)], 1, function(x) sum(x) > 1), 1, 0)
summary_events$any_events <- ifelse(apply(summary_events[, c(3:7)], 1, function(x) sum(x) == 0), 1, 0)
summary_events[summary_events$several_events == 1, 3:7] = 0

events_occ <- summary_events %>%
  group_by(group) %>%
  summarise(total_cases = n(),
            SNV_event_driver_gene = sum(SNV_event_driver_gene ),
            CNV_event  = sum(CNV_event),
            shattered_event  = sum(shattered_event ),
            WGD_event = sum(WGD_event ), 
            Epigenetic_regulatory_gene_event = sum(Epigenetic_regulatory_gene_event),
            several_events  = sum(several_events ),
            any_events = sum(any_events))

events_occ[,3:ncol(events_occ)] <- (events_occ[,3:ncol(events_occ)]/events_occ$total_cases) * 100

events_occ_l  <- events_occ %>% dplyr::select(-total_cases) %>%
  pivot_longer(!group, names_to = "event", values_to = "percentage")

events_occ_l$event <- factor(events_occ_l$event , levels = rev(c("any_events", "SNV_event_driver_gene" , "CNV_event", "shattered_event", "WGD_event", "Epigenetic_regulatory_gene_event", "several_events")))

palette_driver_event <- c("any_events"="#7d7d7d", "SNV_event_driver_gene"="#4EB265", "CNV_event"="#FF9F00", "WGD_event"="#C4DE54", "shattered_event"="#81031D", "Epigenetic_regulatory_gene_event"="#4EC6D0", "several_events"="#DC050C")

ggplot(events_occ_l, aes(x=group, y=percentage, fill=event)) + 
  geom_bar(stat="identity") + theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1, colour = c("Ca A1"="#999933", "Ca A2"="#DDCC77", "Ca B"="#117733", "sc-enriched"="#CC6677")), legend.title =  element_blank()) + 
  xlab("") + ylab("% of samples") + scale_fill_manual(values=palette_driver_event, labels = c("any_events"="None", "SNV_event_driver_gene"="SNV driver", "CNV_event"="CNV driver", "WGD_event"="WGD", "shattered_event"="Shattered regions", "several_events"="Several events", "Epigenetic_regulatory_gene_event"="ER gene variant"))
```















