---
title: "Fig5DE_SupFigs"
author: "E. Mathian and N. Alcala"
date: "`r Sys.Date()`"
output: html_document
---

# Figure 5DE and associated supplementary figures: Whole-slide images deep learning analyses

This document produces Figure 5 from the lungNENomics manuscript, describing the results of Deep-learning analyses of whole-slide images.

## Load libraries

```{r libs, echo=FALSE}
library(tidyverse)
library(ggpubr)
library(RColorBrewer)
#library(dplyr)
#library(tidyr)
#library(optimx)
#library(ggExtra)
#library(webr)
#library(stringr)
#library(tibble)
#library(gridExtra)
#library(reshape2)
#library(tools)
#library(readxl)
#library(scales)
#library(mvtnorm)
#library(RColorBrewer)
#library(irr)  # for Kappa
```

```{r}
setwd("/data/lungNENomics/work/MathianE/pathreview2024/R_Script")
#dir.create("../FiguresStatsPathReview")
```

## Define colors
```{r}
source("/data/lungNENomics/work/Descriptive_manuscript_data/Colour_palettes.R")
yes_no_pal <- c("Yes"="#B2DF8A", "No"="#FB9A99" , "Maybe"="#A6CEE3")
t_cells_size <- c("Unusually small"="#33A02C", "Unusually large"="#E31A1C" , "Medium"="#1F78B4", "NA_T_cells_size"= "#CCCCCC")
nc_ratio_pal <- c("Unusually low"="#33A02C", "Unusually high"="#E31A1C" , "Medium"="#1F78B4", "NA_NC_ratio"= "#CCCCCC")
#"#A6CEE3" "#1F78B4" "#B2DF8A" "#33A02C" "#FB9A99" "#E31A1C" "#FDBF6F" "#FF7F00" "#CAB2D6" "#6A3D9A" "#FFFF99" "#B15928"
```

## import data

We import the list of tiles and their cluster assignments
```{r}
#read_csv("/data/lungNENomics/work/MathianE/pathreview2024/step_2_global_annotation/Partitions_to_annotate_globally.xlsx")
#read_csv("/data/lungNENomics/work/MathianE/pathreview2024/step_2_global_annotation_answer/conacatenated_global_annotation.csv")
Gannot.pre = read_csv("/data/lungNENomics/work/MathianE/pathreview2024/step_2_global_annotation_answer/conacatenated_global_annotation_agg_partition.csv") %>% mutate(partition=as.numeric(str_extract(partition_id,"[0-9]+$")))
load("/data/lungNENomics/work/Descriptive_manuscript_data/Images/Deep_learning_Alexandra/Supervised_Molecular_groups/partition_df_filter2.RData") # list of all tiles in the global partitions

Gextracted_tiles = read_csv("/data/lungNENomics/work/MathianE/LNEN_LCNEC_Molecular_barlow_twin/leinden_highest_attn_bt_proj/FiguresOverviewPerfRoformerLeiden_K250_res2_run3/TilesExtractionRandom50Tiles/ExtractedTiles.csv") # list of tiles actually reviewed by pathologists

#partition_df_filter2
Gannot = left_join(partition_df_filter2 %>% filter(img_id %in% Gextracted_tiles$TilesExtracted) ,Gannot.pre,by=c("partition")) %>% 
  dplyr::rename("patient_id"=sample_id, "erythrocyte"=erythrocytes , "endothelial_cell"="endothelial_cells" , 
                "ciliated_cell"="ciliated_cells", "cartilage_/_chondrocyte"="cartilage", "gobelt_cell"="gobelet_cells", 
                "spindle"="spindle_cells","pyknotic"= "pyknotic_cells","plasmocytoid"="plasmocytoid_cells","polygonal"="polygonal_cells",
                "macrophage"="macrophages","lymphocyte"=lymphocytes,
                "n:c_ratio"="NC_ratio","necrosis"="foci_of_necrosis")

IDmatch = read_csv("/data/lungNENomics/work/MathianE/LNEN_LCNEC_Molecular_barlow_twin/KEY_FILE_tneid_manuscript_id_match.csv")
TBTannot = read_csv("/data/lungNENomics/work/MathianE/pathreview2024/step2_local_annotation_results/Supplementary_S43_TableTileByTileAnnotation.csv")

TBTannot2 = left_join(TBTannot,IDmatch,by=c(patient_id="sample_id"))
colnames(TBTannot2) = str_replace_all(tolower(colnames(TBTannot2)),pattern = " ", replacement = "_")

all_annot = bind_rows(TBTannot2,Gannot)
```

We format the data and compute statistics for each feature
```{r}
all_annot.perSample = all_annot %>% pivot_longer(fibroblast:necrosis,names_to = "Feature") %>% 
  mutate(value_numeric = case_when(value=="No" ~ 0,
                                   value=="Yes" ~ 1,
                                   value=="Unusually small"~0,
                                   value=="Unusually low"~0,
                                   value=="Medium"~0.5,
                                   value=="Unusually high"~1,
                                   value=="Unusually large"~1,
                                   TRUE~NA)) %>% 
  group_by(patient_id,Feature) %>% summarize(value_numeric = mean(value_numeric,na.rm=T),group=archtype_label_combined[1])


all_annot.perSample.stats = data.frame(Feature=NULL,"Statistic"=NULL,"Parameter"=NULL,"P_value"=NULL)
for(feat in unique(all_annot.perSample$Feature)[-23]){ # without tissue architecture, which is not numeric
  print(feat)
  datatmp = all_annot.perSample %>% filter(Feature==feat)
  res = kruskal.test(value_numeric~group,data=datatmp)
  res.means = datatmp %>% group_by(group) %>% summarize(mean = mean(value_numeric,na.rm=T))
  
  means.perm=c()
  for(i in 1:10000){
    group.perm = sample(datatmp$group,replace=F)
    means.perm = rbind(means.perm,c( mean(datatmp$value_numeric[group.perm=="Ca A1"],na.rm=T), 
                    mean(datatmp$value_numeric[group.perm=="Ca A2"],na.rm=T),
                    mean(datatmp$value_numeric[group.perm=="Ca B"],na.rm=T) ) )
  }
  pval_A1A2 = mean( abs(res.means$mean[1]-res.means$mean[2]) <= abs(means.perm[,1]-means.perm[,2]) )
  pval_A1B  = mean( abs(res.means$mean[1]-res.means$mean[3]) <= abs(means.perm[,1]-means.perm[,3]) )
  pval_A2B  = mean( abs(res.means$mean[2]-res.means$mean[3]) <= abs(means.perm[,2]-means.perm[,3]) )
  pval_global  = mean( (abs(res.means$mean[1]-res.means$mean[2])+abs(res.means$mean[1]-res.means$mean[3]))  <= 
          (abs(means.perm[,1]-means.perm[,2]) + abs(means.perm[,1]-means.perm[,3]) ) )
  
  all_annot.perSample.stats = rbind( all_annot.perSample.stats, c(feat,res$statistic,res$parameter,res$p.value,(res.means %>% arrange( group) )$mean,
                                                                  pval_A1A2,pval_A1B,pval_A2B,pval_global) )
}
colnames(all_annot.perSample.stats) = c("Feature","Statistic","Parameter","P_value","Ca A1","Ca A2","Ca B",
                                        "empirical_P_value_A1vsA2","empirical_P_value_A1vsB","empirical_P_value_A2vsB","empirical_P_value_global")
for(j in 2:ncol(all_annot.perSample.stats)) all_annot.perSample.stats[,j] = as.numeric(all_annot.perSample.stats[,j])

all_annot.perSample.stats$q_value = p.adjust(all_annot.perSample.stats$P_value,method = "BH")
all_annot.perSample.stats$empirical_q_value_A1vsA2 = p.adjust(all_annot.perSample.stats$empirical_P_value_A1vsA2,method = "BH")
all_annot.perSample.stats$empirical_q_value_A1vsB  = p.adjust(all_annot.perSample.stats$empirical_P_value_A1vsB,method = "BH")
all_annot.perSample.stats$empirical_q_value_A2vsB  = p.adjust(all_annot.perSample.stats$empirical_P_value_A2vsB,method = "BH")
all_annot.perSample.stats$empirical_q_value_global = p.adjust(all_annot.perSample.stats$empirical_P_value_global,method = "BH")

all_annot.perSample.stats %>% filter(q_value<=0.05 | empirical_q_value<=0.05)

ggplot(all_annot.perSample.stats,aes(x=q_value,y=empirical_q_value)) + geom_point() + 
  geom_smooth(method = "lm")+ geom_hline(yintercept=c(0.05,0.1),linetype="dashed") +
  geom_vline(xintercept=c(0.05,0.1),linetype="dashed") +
  theme_classic()

```


We plot the results
```{r}
FigSDLannottests = ggplot(all_annot.perSample %>% filter(!Feature=="tissue_architecture"),aes(y=value_numeric,x=group,fill=group))+geom_violin(scale = "width") + geom_boxplot(width=0.1,fill="white") + 
  stat_compare_means(
    comparisons=list(
      c("Ca A1", "Ca A2"),c("Ca A1", "Ca B"),c("Ca A2", "Ca B")),
    method="wilcox.test",
    label="p.signif",
    hide.ns=TRUE,
    tip.length=0,label.y.npc = 0,
    step.increase=0.2,
  ) + scale_y_continuous(limits = c(0,1.6),breaks = c(0,0.5,1.0)) + 
  theme_classic() + scale_fill_manual(values=arc4) + facet_wrap(.~Feature) + guides(fill="none")

rownames(all_annot.perSample.stats) = all_annot.perSample.stats$Feature

pheatmap::pheatmap( t((all_annot.perSample.stats%>% filter(empirical_q_value_global<=0.05)) [,c(5:7)]) ,scale = "none")
```