---
title: "CellChat_CCC_spatial_lung"
output: html_document
date: "2025-04-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = "ragg_png")
```

```{r load libraries, echo=TRUE, message=TRUE}
library(Seurat)
library(ggplot2)
library(dplyr)
library(IRIS)
library(CellChat)
library(reshape)
library(circlize)

```

# Part I: Data input & processing and initialization of CellChat object

## Load data and concatenate labels from IRIS predictions

```{r load data}

#setwd("Z:/Training-MG/files/data/Project1_spatial_lnets/data")

lung_A <-readRDS("Z:/Training-MG/files/data/Project1_spatial_lnets/data/OneDrive_1_29-11-2024/adata_LNEN071-IARC-A_scanpy_06052024.rds")

lung_B <-readRDS("Z:/Training-MG/files/data/Project1_spatial_lnets/data/OneDrive_1_29-11-2024/adata_LNEN084-IARC-B_scanpy_06052024.rds")

lung_C <-readRDS("Z:/Training-MG/files/data/Project1_spatial_lnets/data/OneDrive_1_29-11-2024/adata_LNEN107-IARC-C_scanpy_06052024.rds")

lung_D <-readRDS("Z:/Training-MG/files/data/Project1_spatial_lnets/data/OneDrive_1_29-11-2024/adata_LNEN206-IARC-D_scanpy_06052024.rds")

load("Z:/Training-MG/files/data/Project1_spatial_lnets/data/IRIS_object.Rdata")
```
### extract IRIS spatial domains
```{r save IRIS annotations to file}
path_iris_annot <- "C:/Users/limy/OneDrive - International Agency for Research on Cancer/Documents/Cell communication/LungNENomics/code/data/iris_annotations1.csv"

write.csv(IRIS_object@spatialDomain, path_iris_annot, row.names = FALSE)
```

```{r save IRIS proportions to file}
path_iris_annot <- "C:/Users/limy/OneDrive - International Agency for Research on Cancer/Documents/Cell communication/LungNENomics/code/data/IRIS_proportions_all.csv"

write.csv(IRIS_object@IRIS_Prop, path_iris_annot, row.names = FALSE)
```


```{r extract IRIS spatial domains}
iris_slice1 <- IRIS_object@spatialDomain[IRIS_object@spatialDomain$Slice == "Slice1",]
iris_slice2 <- IRIS_object@spatialDomain[IRIS_object@spatialDomain$Slice == "Slice2",]
iris_slice3 <- IRIS_object@spatialDomain[IRIS_object@spatialDomain$Slice == "Slice3",]
iris_slice4 <- IRIS_object@spatialDomain[IRIS_object@spatialDomain$Slice == "Slice4",]

# change rownames by spot names

rownames(iris_slice1) <- iris_slice1$spotName
rownames(iris_slice2) <- iris_slice2$spotName
rownames(iris_slice3) <- iris_slice3$spotName
rownames(iris_slice4) <- iris_slice4$spotName

```
### Add Iris domains annotation
```{r add Iris domains annotation}
lung_A <- AddMetaData(lung_A, iris_slice1)
lung_B <- AddMetaData(lung_B, iris_slice2)
lung_C <- AddMetaData(lung_C, iris_slice3)
lung_D <- AddMetaData(lung_D, iris_slice4)
```


### Visualise number of spots in each domain
```{r}
plot_function <- function(iris_slice)
{
slice_counts <- as.data.frame(table(iris_slice$IRIS_domain))
colnames(slice_counts) <- c("Domain", "Count")  # Rename columns for clarity

# Visualize with ggplot2
ggplot(slice_counts, aes(x = Domain, y = Count)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = paste("Domain counts for", iris_slice$Slice[1]), x = "Domain", y = "Count") +
  theme(
    plot.title = element_text(size=22, hjust=0.5),
    axis.title.x = element_text(size=18),
    axis.title.y = element_text(size=18),
    axis.text.x = element_text(size=14),
    axis.text.y = element_text(size=14)
  )
#ggsave(paste("domain_counts_",iris_slice$Slice[1],".png"), width=7, height=5)
}
```


```{r visualise number of spots in each domain}
plot_function(iris_slice1)
```


## Creat CellChat object

### Prepare input for CellChat

```{r}
make_cellchat_input <- function(lung_object)
{
    # normalize data
norm_data <- normalizeData(lung_object[["RNA"]]@data)
head(norm_data)

# normalized count matrix
data.input <- norm_data

# cell/spot annotations
labels_iris <- lung_object$IRIS_domain
labels_iris[labels_iris == "0"] <- "00"
#creating metadata for cellchat
meta_iris <- data.frame(labels = labels_iris, row.names = names(labels_iris))
head(meta_iris)

# spatial coordinates
spatial.locs <- data.matrix(lung_object@meta.data[,c("x","y")])

# scale factors
spot.diameter.fullres = lung_object@images$slice1@scale.factors$spot
spot.size = 65 # the theoretical spot size (um) in 10X Visium
conversion.factor = spot.size/spot.diameter.fullres
print(paste("conversion factor", conversion.factor))

# spatial factors parameter: ratio (conversion coefficient from pixels to microns) and tol(tolerance factor)
spatial.factors = list(ratio = conversion.factor, tol = spot.size/2)

# check consistency
d.spatial <- computeCellDistance(coordinates = spatial.locs, ratio = spatial.factors$ratio, tol = spatial.factors$tol)
min(d.spatial[d.spatial!=0]) # this value should approximately equal 100um for 10X Visium data


return(createCellChat(object = data.input, meta = meta_iris, group.by = "labels", datatype = "spatial", coordinates = spatial.locs, spatial.factors = spatial.factors)) 

}
  
```


```{r create CellChat object from seurat sample}

cellchat_A <- make_cellchat_input(lung_A)
cellchat_B <- make_cellchat_input(lung_B)
cellchat_C <- make_cellchat_input(lung_C)
cellchat_D <- make_cellchat_input(lung_D)

```


### Set LR interaction database
```{r set LR interaction database}
CellChatDB <- CellChatDB.human 
showDatabaseCategory(CellChatDB)

# Show the structure of the database
dplyr::glimpse(CellChatDB$interaction)
```

### Subset Secreted Signaling and Cell-cell contact interations
```{r subset all except non-protein signaling}
CellChatDB.use <- subsetDB(CellChatDB)
#CellChatDB.use <- subsetDB(CellChatDB, search = c("Secreted Signaling","Cell-Cell Contact"))
#CellChatDB.use <- subsetDB(CellChatDB, search = "ECM-Receptor")

# Show the structure of the database
showDatabaseCategory(CellChatDB.use)

# set the used database in the object
cellchat_A@DB <- CellChatDB.use  
cellchat_B@DB <- CellChatDB.use  
cellchat_C@DB <- CellChatDB.use  
cellchat_D@DB <- CellChatDB.use  
```

```{r function to process cellchat object}
process_cellchat <- function(cellchat)
{
  # subset the expression data of signaling genes for saving computation cost
  cellchat <- subsetData(cellchat) ## save in cellchat@data.signaling
  cellchat <- identifyOverExpressedGenes(cellchat)
  cellchat <- identifyOverExpressedInteractions(cellchat)
  
  return(cellchat)

}
```

```{r}
cellchat_A <- process_cellchat(cellchat_A)
cellchat_B <- process_cellchat(cellchat_B)
cellchat_C <- process_cellchat(cellchat_C)
cellchat_D <- process_cellchat(cellchat_D)
```
# Part II:Inference of cell-cell interaction network

## Compute the communication probability and infer cellular communication network
```{r function to infer a network}
make_network <- function(cellchat)
{
  # use truncated mean 25%
  cellchat <- computeCommunProb(cellchat, type = "triMean", distance.use = TRUE, contact.range=100)
  
  # Filter out the cell-cell communication if there are only few number of cells in certain cell groups
  cellchat <- filterCommunication(cellchat, min.cells = 10)
  
  # infer CCI at a signaling pathway level
  cellchat <- computeCommunProbPathway(cellchat)
 
  
  return(cellchat)
}
```


```{r}
cellchat_A <- make_network(cellchat_A)
cellchat_B <- make_network(cellchat_B)
cellchat_C <- make_network(cellchat_C)
cellchat_D <- make_network(cellchat_D)

#Extract the computed network in a dataframe
network_A <- subsetCommunication(cellchat_A)
network_B <- subsetCommunication(cellchat_B)
network_C <- subsetCommunication(cellchat_C)
network_D <- subsetCommunication(cellchat_D)
```

```{r save cellchat objects}
saveRDS(cellchat_A, file = "CellChat objects/cellchat_A.rds")
saveRDS(cellchat_B, file = "CellChat objects/cellchat_B.rds")
saveRDS(cellchat_C, file = "CellChat objects/cellchat_C.rds")
saveRDS(cellchat_D, file = "CellChat objects/cellchat_D.rds")
```

