---
title: "ExtDataFig4D"
author: "N. Alcala"
date: "`r Sys.Date()`"
output: html_document
---

# Extended Data Figure 4 panel D: Proliferation

This document produces panel D from Extended Data Figure 4 from the lungNENomics manuscript, describing cellular proliferation across the groups.

## Panel D

### Load libraries

```{r pressure, echo=FALSE}
library(tidyverse)
library(patchwork)
library(ggbeeswarm)
```

### load lungNENomics color schemes
For now, this only runs locally. Nevertheless, not necessary to run, just provides cosmetics.
```{r}
source("/data/lungNENomics/work/Descriptive_manuscript_data/Colour_palettes.R")
colors_clusters = c("Ca A1"="#999933", "Ca A2"= "#DDCC77", "Ca B"= "#117733", "sc-enriched"="#CC6677",  
                    "Typical"="#9DB802","Atypical"="#025B0E","undetermined"="#CEDB80",
                    "LCNEC"="#824833" ,"SCLC"="#000000",
                    SINET = "#2ad4ffff") #, "Adeno-Ca"="#ff80e5ff", "Squamous-Cell-Ca"="#8080ffff") 
```


### load metadata
We need to merge the metadata for the LCNEC and SCLC cohorts (data harmonized in Gabriel et al. 2020) with the clinical data from the lungNENomics cohort (in Table S1)

```{r}
Attributes = read_tsv(unzip('../../Data/GabrielEtAl2020/Attributes.txt.zip')[1])

load("/data/lungNENomics/work/Descriptive_manuscript_data/MOFA/MOFA.Exp.Meth.Alt.CNV_lungNENomicsCombined/ParetoTI/variables_archetypes_MOFA_LNET.RData") # data also contained in Table S1
gigascience_attributes = read_tsv("/data/lungNENomics/work/SextonoatesA/Attributes.txt")
var_LNET$sample_id = as.character(var_LNET$sample_id)
var_LNET$group = var_LNET$archetype_k4_LF3_label

var_cli = bind_rows(var_LNET, Attributes %>% filter(Histopathology_simplified%in%c("LCNEC","SCLC")) %>% 
                      mutate(sample_id = Sample_ID, group=Histopathology_simplified) %>% dplyr::select(sample_id,group) )
var_cli = bind_rows(var_cli,tibble(sample_id=c("LCNEC3T","LCNEC4T"),group="LCNEC") )
var_cli$type[is.na(var_cli$type)] = var_cli$group[is.na(var_cli$type)]
```

### load expression data

```{r}
load("/data/lungNENomics/work/Descriptive_manuscript_data/RNA_sequencing_data/gene_TPM_nosex_matrix_PCA_LCNEC_SCLC_ITH_TR.RData") # expression
load("/data/lungNENomics/work/Descriptive_manuscript_data/MOFA/MOFA.Exp.Meth.Alt.CNV_lungNENomicsCombined/ParetoTI/variables_archetypes_MOFA_LNET.RData")
ki67.vst <- read.csv(file="/data/lungNENomics/work/Descriptive_manuscript_data/RNA_sequencing_data/vstexpr_MKI67_PCA_LCNEC_SCLC_ITH_TR.csv", row=1)
load("/data/lungNENomics/work/Descriptive_manuscript_data/gencode_hg38_v33.RData") # annotation

dim(TPM_nosex)
head(var_LNET)
table(var_LNET$sample_id, var_LNET$archetype_k4_LF3_label)

TPM_nosex = TPM_nosex[,!colnames(TPM_nosex)%in%c("LNEN246_TU","LNEN251_TU1")]

ki67.tpm = bind_cols( t(TPM_nosex["ENSG00000148773.14",]) ,sample_id =  colnames(TPM_nosex))
```

We merge the tables and exclude samples without matching RNA-seq
```{r}
expr_MKI67_tpm_vst = left_join(ki67.tpm, left_join(ki67.vst,
                          var_LNET[is.na(var_LNET$RNAseq_batch)==F & is.na(var_LNET$archetype_k4_LF3_label)==F,]  %>% 
                            dplyr::select(sample_id, archetype_k4_LF3_label,type), by="sample_id") )#


expr_MKI67_tpm_vst = expr_MKI67_tpm_vst %>% filter(!sample_id %in%c("LNEN246_TU","LNEN251_TU1") )

# check merge
which( is.na(expr_MKI67_tpm_vst$ENSG00000148773.14) )
which( is.na(expr_MKI67_tpm_vst$MKI67_expression) )

colnames(expr_MKI67_tpm_vst)[1] = "MKI67_expression_TPM"
colnames(expr_MKI67_tpm_vst)[3] = "MKI67_expression_vst"
```

### Statistical tests

We use vst expression to perform statistical tests, as it is closest to normal
```{r}
summary(aov(MKI67_expression_vst ~ archetype_k4_LF3_label, data = expr_MKI67_tpm_vst))[[1]][1,5] # p-value: 9.403244e-10

t.test(expr_MKI67_tpm_vst$MKI67_expression_vst[which(expr_MKI67_tpm_vst$archetype_k4_LF3_label %in% c("Ca A1", "Ca A2"))] ~ 
         expr_MKI67_tpm_vst$archetype_k4_LF3_label[which(expr_MKI67_tpm_vst$archetype_k4_LF3_label %in% c("Ca A1", "Ca A2"))], alternative = "two.sided") # p-value = 0.03479 *
t.test(expr_MKI67_tpm_vst$MKI67_expression_vst[which(expr_MKI67_tpm_vst$archetype_k4_LF3_label %in% c("Ca A1", "Ca B"))] ~ 
         expr_MKI67_tpm_vst$archetype_k4_LF3_label[which(expr_MKI67_tpm_vst$archetype_k4_LF3_label %in% c("Ca A1", "Ca B"))], alternative = "two.sided") # p-value = 0.7354
t.test(expr_MKI67_tpm_vst$MKI67_expression_vst[which(expr_MKI67_tpm_vst$archetype_k4_LF3_label %in% c("Ca A1", "sc-enriched"))] ~ 
         expr_MKI67_tpm_vst$archetype_k4_LF3_label[which(expr_MKI67_tpm_vst$archetype_k4_LF3_label %in% c("Ca A1", "sc-enriched"))], alternative = "two.sided") # p-value = 7.331e-07 ***


t.test(expr_MKI67_tpm_vst$MKI67_expression_vst[which(expr_MKI67_tpm_vst$archetype_k4_LF3_label %in% c("Ca A2", "Ca B"))] ~ 
         expr_MKI67_tpm_vst$archetype_k4_LF3_label[which(expr_MKI67_tpm_vst$archetype_k4_LF3_label %in% c("Ca A2", "Ca B"))], alternative = "two.sided") # p-value = 0.2068
t.test(expr_MKI67_tpm_vst$MKI67_expression_vst[which(expr_MKI67_tpm_vst$archetype_k4_LF3_label %in% c("Ca A2", "sc-enriched"))] ~ 
         expr_MKI67_tpm_vst$archetype_k4_LF3_label[which(expr_MKI67_tpm_vst$archetype_k4_LF3_label %in% c("Ca A2", "sc-enriched"))], alternative = "two.sided") # p-value = 6.232e-08 ***

t.test(expr_MKI67_tpm_vst$MKI67_expression_vst[which(expr_MKI67_tpm_vst$archetype_k4_LF3_label %in% c("Ca B", "sc-enriched"))] ~ 
         expr_MKI67_tpm_vst$archetype_k4_LF3_label[which(expr_MKI67_tpm_vst$archetype_k4_LF3_label %in% c("Ca B", "sc-enriched"))], alternative = "two.sided") # p-value = 1.856e-07 ***

# by histological type
summary(aov(MKI67_expression_vst ~ type, data = expr_MKI67_tpm_vst))[[1]][1,5] # p-value: 1.263021e-16


t.test(expr_MKI67_tpm_vst$MKI67_expression_vst[which(expr_MKI67_tpm_vst$type %in% c("Typical", "Atypical"))] ~ 
         expr_MKI67_tpm_vst$type[which(expr_MKI67_tpm_vst$type %in% c("Typical", "Atypical"))], alternative = "two.sided")$p.value # p-value = 8.493567e-17 ***
t.test(expr_MKI67_tpm_vst$MKI67_expression_vst[which(expr_MKI67_tpm_vst$type %in% c("Typical", "Carcinoid"))] ~ 
         expr_MKI67_tpm_vst$type[which(expr_MKI67_tpm_vst$type %in% c("Typical", "Carcinoid"))], alternative = "two.sided") # p-value = 0.04355 *
t.test(expr_MKI67_tpm_vst$MKI67_expression_vst[which(expr_MKI67_tpm_vst$type %in% c("Typical", "NET G3"))] ~ 
         expr_MKI67_tpm_vst$type[which(expr_MKI67_tpm_vst$type %in% c("Typical", "NET G3"))], alternative = "two.sided") # p-value = 0.4254

t.test(expr_MKI67_tpm_vst$MKI67_expression_vst[which(expr_MKI67_tpm_vst$type %in% c("Atypical", "Carcinoid"))] ~ 
         expr_MKI67_tpm_vst$type[which(expr_MKI67_tpm_vst$type %in% c("Atypical", "Carcinoid"))], alternative = "two.sided") # p-value = 6.268e-05 ***
t.test(expr_MKI67_tpm_vst$MKI67_expression_vst[which(expr_MKI67_tpm_vst$type %in% c("Atypical", "NET G3"))] ~ 
         expr_MKI67_tpm_vst$type[which(expr_MKI67_tpm_vst$type %in% c("Atypical", "NET G3"))], alternative = "two.sided") # p-value = 0.7738

t.test(expr_MKI67_tpm_vst$MKI67_expression_vst[which(expr_MKI67_tpm_vst$type %in% c("Carcinoid", "NET G3"))] ~ 
         expr_MKI67_tpm_vst$type[which(expr_MKI67_tpm_vst$type %in% c("Carcinoid", "NET G3"))], alternative = "two.sided") # p-value = 0.5014 

mean(expr_MKI67_tpm_vst$MKI67_expression_vst[which(expr_MKI67_tpm_vst$type=="Typical")], na.rm=TRUE) # 7.999174
mean(expr_MKI67_tpm_vst$MKI67_expression_vst[which(expr_MKI67_tpm_vst$type=="Atypical")], na.rm=TRUE) # 9.875885
mean(expr_MKI67_tpm_vst$MKI67_expression_vst[which(expr_MKI67_tpm_vst$type=="Carcinoid")], na.rm=TRUE) # 8.575349
mean(expr_MKI67_tpm_vst$MKI67_expression_vst[which(expr_MKI67_tpm_vst$type=="NET G3")], na.rm=TRUE) # 10.65458
```

### Plots
```{r}
ylabpos = 1.4
EDF4Dleft <- ggplot(data=expr_MKI67_tpm_vst %>% filter(!is.na(archetype_k4_LF3_label)), 
             aes(x=archetype_k4_LF3_label, y = log10(MKI67_expression_TPM+1) )) +
  geom_violin(aes(fill=archetype_k4_LF3_label)) + 
  geom_boxplot(fill="white", width=0.1, outlier.size = 0.4) +
  scale_fill_manual(name = "Archetype", values=c("Ca A1"="#999933", "Ca A2"="#DDCC77", "Ca B"="#117733", "sc-enriched"="#CC6677")) +
  theme_classic() + ylab("MKI67 expression [log10(TPM+1)]") + xlab("Archetype") +
  theme(legend.position = "right") +
  scale_y_continuous(breaks = c(0,0.5,1,1.5,2), limits=c(0,2.5)) +
  guides(color=guide_legend(nrow=1,byrow=TRUE)) + 
  geom_segment(aes(x=1,xend=2,y=ylabpos,yend=ylabpos)) +  geom_text(aes(label = "*", y = ylabpos+0.1, x=1.5)) +
  geom_segment(aes(x=1,xend=4,y=ylabpos+.3,yend=ylabpos+0.3)) +  geom_text(aes(label = "***", y = ylabpos+0.3+0.1, x=2.5)) +
  geom_segment(aes(x=2,xend=4,y=ylabpos+0.3*2,yend=ylabpos+0.3*2)) +  geom_text(aes(label = "***", y = ylabpos+0.3*2+0.1, x=3)) +
  geom_segment(aes(x=3,xend=4,y=ylabpos+0.3*3,yend=ylabpos+0.3*3)) +  geom_text(aes(label = "***", y = ylabpos+0.3*3+0.1, x=3.5)) +
  #geom_text(aes(label = "n=109", y = 1, x=1),color="black",size=5) + 
  #geom_text(aes(label = "n=89", y = 1, x=2),color="black",size=5) + 
  #geom_text(aes(label = "n=62", y = 1, x=3),color="black",size=5) + 
  #geom_text(aes(label = "n=13", y = 1, x=4),color="black",size=5) +
  theme(axis.text.y=element_text(size=11), axis.text.x=element_text(size=11), 
        axis.title.x=element_text(size=11), axis.title.y=element_text(size=11)) + 
  theme(legend.text=element_text(size=11), legend.title=element_text(size=11)) +
  labs(title="MKI67 expression by group") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Molecular group") + guides(fill="none")

ggsave("/data/lungNENomics/work/alcalan/Figures/ExtendedDataFig_4D_MKI67.svg",plot = EDF4Dleft,height=2.5,width=2.5)
```